Fig2: Candidate genes

### 2020-05-20 analysis/figure-1.Rmd
music removed, correlations moved to suppl, fig2 needs more

suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
  arrange( tissue, sex, direction, gene)
head(suppltable2)
write_csv(suppltable2, "../results/suppltable1.csv")

>>>>

suppltable2 <- rbind(DEGcontrol, DEGbldg)  %>%
  arrange( tissue, sex, direction, gene)
head(suppltable2)
write_csv(suppltable2, "../results/suppltable2.csv")

### 2020-05-20 analysis/figure-2.Rmd
music removed, correlations moved to suppl, fig2 needs more

@@

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, control_bldg, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, control_bldg:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
table1$numDEGs <- rowSums(is.na(table1)) # count NAs to know how many are NS

table1 <- table1 %>% 
  mutate(sig = ifelse(numDEGs == 6, "NS", "DEG")) %>% 
  arrange(gene)  %>%  select(-sig, -numDEGs)
table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
head(table1)

kable(table1)
```

## Candidate Correlations

```{r}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv")
head(candidatevsd)

curleychampagnegenes <- curleychampagnegenes %>% pull(gene)

hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)

@@

hyp3 <- makecorrdf(sexlevels, "hypothalamus", curleychampagnegenes)  
pit3 <- makecorrdf(sexlevels, "pituitary", curleychampagnegenes)  
gon3 <- makecorrdf(sexlevels, "gonad", curleychampagnegenes)

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad")

@@

b1 <- plotcorrplot(hyp3, "females and males") + theme(legend.position = "none")  + labs(y = "Hypothalamus", title = "Candidate gene correlations")
b2 <- plotcorrplot(pit3, NULL) + theme(legend.position = "none") + labs(y = "Pituitary")
b3 <- plotcorrplot(gon3, NULL) + theme(legend.position = "none")  + labs(y = "Gonad")

c1 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
c5 <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)

c2 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
c6 <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL, y = NULL)

d1 <- candidateboxplot("hypothalamus", c("DRD1"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
d3 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL )  + theme( strip.text = element_blank())
d4 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = NULL  )  + theme(strip.text = element_blank())
d5 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL) + theme(strip.text = element_blank())
d6 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())

b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))

bcd <- plot_grid(b,c,d, ncol = 3, rel_widths = c(1,1,1), labels = c("A", "B", "C"), label_size = 8)

@@

```{r fig3, fig.height=3, fig.width=7.25}

musicplot <- function(whichgene, whichtissue){

    p <- candidatevsd %>%
      group_by(treatment, tissue, gene, sex)  %>% 
      summarize(median = median(counts, na.rm = T), 
                se = sd(counts,  na.rm = T)/sqrt(length(counts))) %>%
      dplyr::mutate(scaled = rescale(median, to = c(0, 7))) %>%
      dplyr::mutate(image = "../figures/images/musicnote.png")   %>%
      filter(
        gene %in% whichgene,
        tissue %in% whichtissue      ) %>% 
      collect() %>%
      drop_na() %>%
      ggplot( aes(x = treatment, y = median)) +
      geom_errorbar(aes(ymin = median - se, 
                        ymax = median + se, color = "white"),  width=0) +
      geom_image(aes(image=image), size = 0.1)+
      theme_B3() +
      theme(legend.position = "none",
            title = element_text(face = "italic"),
            strip.text = element_text(color = "white"),
            axis.text.x = element_text(angle = 45, hjust = 1),
            axis.text.y = element_blank(),
            axis.line = element_blank(), axis.ticks = element_blank()) +
      scale_color_manual(values = allcolors) +
      labs(y = " ",x = NULL ) +
      facet_wrap(~sex, nrow = 2, scales = "free_y")
      return(p)
}

music <- png::readPNG("../figures/images/fig_music.png")
music <- ggdraw() +  draw_image(music, scale = 1)

musicalgenes <- png::readPNG("../figures/images/fig_musicalgenes.png")
musicalgenes <- ggdraw() +  draw_image(musicalgenes, scale = 1)

p1 <- musicplot(c("AVP", "OXT"), "hypothalamus") + labs(subtitle = "Hypothalamic AVP & OXT" ) + theme(axis.text.x = element_blank())
p2 <- musicplot(c("AVP", "OXT"), "pituitary") + labs(subtitle = "Pituitary AVP & OXT " ) + theme(axis.text.x = element_blank())
p3 <- musicplot(c("AVP", "OXT"), "gonad") + labs(subtitle = "Gonadal  AVP & OXT"  )

p123 <- plot_grid(music, p1, music, p3, ncol = 2, rel_widths = c(0.125,1))

plot_grid(musicalgenes, p123, nrow = 1, rel_widths = c(1.5,1))

```{r writefiles}
#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
#write.csv(candidatevsd, "../results/candidatecounts.csv")
write.csv(table1, "../results/table1.csv")
write.csv(candidategenes, "../results/candidategenes.csv")
```

>>>>

## Candidate genes

```{r candidategenes}
parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
head(parentalcaregenes)

curleychampagnegenes <- parentalcaregenes %>% distinct(literature) %>% drop_na() %>% pull(literature)
GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
candidategenes <- parentalcaregenes %>% pull(gene) 
candidategenes
```

@@

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
table1[is.na(table1)] <- " " # replace NA with blank space so it's pretty
head(table1)

## Candidate VSDs

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels)
tail(candidatevsd)

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad") 
```

## Candidate Correlations - SUppl fig 1

```{r correlations}
hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)

@@

hyp3 <- makecorrdf("female", "hypothalamus", GOgenes)  
pit3 <- makecorrdf("female", "pituitary", GOgenes)  
gon3 <- makecorrdf("female", "gonad", GOgenes)

hyp4 <- makecorrdf("male", "hypothalamus", GOgenes)  
pit4 <- makecorrdf("male", "pituitary", GOgenes)  
gon4 <- makecorrdf("male", "gonad", GOgenes)  
```

```{r supplfig-2, fig.width=7.25, fig.height=7.25}

b1 <- plotcorrplot(hyp1, "females") + labs(y = "Hypothalamus", title = "Parental Care Literature")  + 
  scale_x_discrete(position = "top")   + theme(axis.text.x = element_text(vjust = -0.25))
b2 <- plotcorrplot(pit1, NULL)  + labs(y = "Pituitary") + 
  theme(axis.text.x = element_blank())
b3 <- plotcorrplot(gon1, NULL)  + labs(y = "Gonad") + theme(axis.text.x = element_text(vjust = 0.25))

b4 <- plotcorrplot(hyp2, "males") + labs( title =  " ") + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = -0.25)) +
  scale_x_discrete(position = "top") 
b5 <- plotcorrplot(pit2, NULL)    + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank())
b6 <- plotcorrplot(gon2, NULL)  + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

b7 <- plotcorrplot(hyp3, "females") + labs( title = "Parental Care GO") + 
   scale_x_discrete(position = "top") + theme(axis.text.x = element_text(vjust = -0.25))
b8 <- plotcorrplot(pit3, NULL)   + 
  theme(axis.text.x = element_blank())
b9 <- plotcorrplot(gon3, NULL)  + theme(axis.text.x = element_text(vjust = 0.25))

b10 <- plotcorrplot(hyp4, "males") + labs( title =  " ") + 
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(vjust = -0.25)) + scale_x_discrete(position = "top") 
b11 <- plotcorrplot(pit4, NULL)    + 
  theme(axis.text.x = element_blank() , axis.text.y = element_blank())
b12 <- plotcorrplot(gon4, NULL)   + theme( axis.text.y = element_blank(), axis.text.x = element_text(vjust = 0.25))

forlegend <- plotcorrplot(gon4, NULL) + theme(legend.position = "bottom") 
mylegend <- get_legend(forlegend)

allcorplots <- plot_grid(b1,b4,b7,b10,
          b2,b5,b8,b11,
          b3,b6,b9,b12, rel_heights = c(1.4,1,1.2),
          labels = c("A", " ", "B"), label_size = 8, rel_widths = c(1.4,1.2,1.2,1))

supplfig2 <- plot_grid(allcorplots, mylegend, ncol = 1, rel_heights = c(1,0.1))
supplfig2

pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
plot(supplfig2)
dev.off()
```

@@

c1 <- scattercorrelations(FH, FH$DRD4, FH$HTR2C, "#969696", "DRD4", "HTR2C") + labs(subtitle = "female", title =  "Top correlation")  
c3 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = NULL) 
c5 <- scattercorrelations(FG, FG$ESR1, FG$PGR, "#969696", "ESR1", "PGR")   + labs(subtitle = NULL)

c2 <- scattercorrelations(MH, MH$DRD4, MH$HTR2C, "#525252", "DRD4", "HTR2C")  + labs(subtitle = "male", title = "", y = NULL)  
c4 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1")  + labs(subtitle = NULL, y = NULL)  
c6 <- scattercorrelations(MG, MG$ESR1, MG$PGR, "#525252", "ESR1", "PGR")  + labs(subtitle = NULL, y = NULL)

d1 <- candidateboxplot("hypothalamus", c("DRD4"), sexlevels) + labs(x = NULL, title = "Temporal expression pattern" )  
d2 <- candidateboxplot("hypothalamus", c("HTR2C"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())
d3 <- candidateboxplot("pituitary", c("AVPR1A"), sexlevels) + labs(x = NULL )  + theme( strip.text = element_blank())
d4 <- candidateboxplot("pituitary", c("CRHR1"), sexlevels) + labs(x = NULL  )  + theme(strip.text = element_blank())
d5 <- candidateboxplot("gonad", c("ESR1"), sexlevels) + labs(x = NULL) + theme(strip.text = element_blank())
d6 <- candidateboxplot("gonad", c("PGR"), sexlevels) + labs(x = NULL ) + theme(strip.text = element_blank())

#b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))

bcd <- plot_grid(c,d, ncol = 2, rel_widths = c(1,1), labels = c("A", "B"), label_size = 8)

@@

```{r writefiles}
#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
#write.csv(candidatevsd, "../results/candidatecounts.csv")
write.csv(table1, "../results/table1.csv")
```
